---
title: It's Getting Hot in Here
subtitle: Analysis of Global CO2 Emissions Over Time
author: Impressive Squirtle <br> Cassidy Cheesman, Joanne Wang, Ming DeMers, Elisabeth Pan, and Ivan Lin
date: 5/5/23
format: revealjs
editor: visual
execute:
  echo: false
---

## Introduce the topic and motivation

-   Climate change is the biggest challenge facing our generation, with increased global CO2 emissions being one of the leading contributors to this issue.

-   In order to combat the effects of climate change, we must first understand the data behind it.

-   Research questions:

    -   What is the relationship between GDP and CO2 emissions for developing versus developed countries in 2018?

    -   Is the true proportion of countries with a CO2 emission greater than the US in the twentieth century different from those in the twenty-first century?

## Introduce the data

-   Created to show how emissions are changing in each country

-   The Our World in Data team created the original dataset on behalf of the Our World in Data project

-   The instances that comprise the dataset represent countries (per year from 1992 to 2018) with 2484 instances total.

-   The observations (rows) are countries per year and the attributes (columns) are various emission data for each country per year

-   The current dataset we will use is a subset of the original dataset, with all the tonnes units converted into kilotons and the variable names altered

## Highlights from EDA

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
library(skimr)
library(rcis)
```

```{r}
#| label: read-csv-glimpse

pollution <- read_csv("data/global_emissions.csv")
```

```{r}
#| label: gdp-developed
#visualizing change in GDP over time for a developed country
pollution_usa <- pollution |>
  filter(Country.Name == "United States")

ggplot(data = pollution_usa, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in USA", 
    x = "Year",
    y = "GDP (international$)"
  )
```

```{r}
#| label: gdp-developing
#visualizing change in GDP over time for a developing country
pollution_pakistan <- pollution |>
  filter(Country.Name == "Pakistan")

ggplot(data = pollution_pakistan, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Pakistan", 
    x = "Year",
    y = "GDP (international$)"
  ) 
```

```{r}
#| label: CO2-developed
#visualizing change in CO2 emissions over time for a developed country
ggplot(data = pollution_usa, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in USA", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```

```{r}
#| label: CO2-developing
#visualizing change in CO2 emissions over time for a developing country
ggplot(data = pollution_pakistan, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Pakistan", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```

```{r}
#| label: gdp-CO2-1992
#visualizing relationship between GDP and total CO2 emissions in 1992
pollution_92 <- pollution |>
  filter(Year == 1992)

ggplot(data = pollution_92,
       mapping = aes(x = Country.GDP, y = Emissions.Production.CO2.Total)
       ) + 
  geom_point(alpha = 0.5) + 
  xlim(0, 3e+12) + 
  ylim(0, 2000) + 
  theme_minimal() + 
  labs(
    title = "Total CO2 Emissions vs. Country GDP in 1992", 
    x = "GDP (international$)",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```

```{r}
#| label: gdp-CO2-2018
#visualizing relationship between GDP and total CO2 emissions in 2018
pollution_18 <- pollution |>
  filter(Year == 2018)

ggplot(data = pollution_18,
       mapping = aes(x = Country.GDP, y = Emissions.Production.CO2.Total)
       ) + 
  geom_point(alpha = 0.5) + 
  xlim(0, 4e+12) + 
  ylim(0, 2000) + 
  theme_minimal() + 
  labs(
    title = "Total CO2 Emissions vs. Country GDP in 2018", 
    x = "GDP (international$)",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```

## Inference/modeling/other analysis

Your results here

## Inference/modeling/other analysis

Your results here

## Conclusions + future work
