---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
library(skimr)
library(tidymodels)
library(openintro)
```

```{r}
#| label: read-csv-glimpse

pollution <- read_csv("data/global_emissions.csv")
glimpse(pollution)

```

# Introduction

One cannot think of global climate change and environmental sustainability without thinking of CO2. According to NASA, "Carbon dioxide in the atmosphere warms the planet, causing climate change. Human activities have raised the atmosphere's carbon dioxide content by 50% in less than 200 years," and according to Climate.gov, " By adding more carbon dioxide to the atmosphere, people are supercharging the natural greenhouse effect, causing global temperature to rise." In our analysis, we intend to explore how global CO2 emissions correlate with the GDP growth rate for developing versus developed countries. Additionally, we aim to investigate the differences in CO2 production between the first year data was available (1992) and the most recent year data was available (2018) for a certain country.

# Data description

### Motivation

-   **For what purpose was the dataset created?**

    -   The dataset was created to show how emissions are changing in each country and allows for comparisons of various countries and regions in relation to overall emission trends along with their sources.

-   **Who created the dataset (e.g., which team, research group) and on behalf of which entity (e.g., company, institution, organization)?**

    -   The Our World in Data team created the original dataset on behalf of the Our World in Data project, which is a project of the Global Change Data Lab (a non-profit organization based in the United Kingdom), and the current dataset we will be using is a subset from the original dataset from the CORGIS Dataset Project created by Sam Donald.

-   **Who funded the creation of the dataset?**

    -   The Global Change Data Lab funded the creation of the dataset and it is a registered charity in England and Wales. Their work is made possible by donations of individuals and organizations that have made contributions to support their project.

### Composition

-   **What do the instances that comprise the dataset represent (e.g., documents, photos, people, countries)?**

    -   The instances that comprise the dataset represent countries (per year from 1992 to 2018).

-   **How many instances are there in total (of each type, if appropriate)?**

    -   There are 2484 instances total.

-   **What are the observations (rows) and the attributes (columns)?**

    -   The observations (rows) are countries per year.

    -   The attributes (columns) are year, country name, country code, country GDP, country population, emissions production CH4, emissions production N2O, emissions production CO2 cement, emissions production CO2 coal, emissions production CO2 gas, emissions production CO2 oil, emissions production CO2 flaring, emissions production CO2 other, emissions production CO2 total, emissions global share CO2 cement, emissions global share CO2 coal, emissions global share CO2 gas, emissions global share CO2 oil, emissions global share CO2 flaring, emissions global share CO2 total.

### Collection Process

-   **How was the data associated with each instance acquired?**

    -   The original dataset is built upon a number of datasets and processing steps. The datasets that were processed include Statistical review of world energy (BP), International energy data (EIA), Primary energy consumption (Our World in Data based on BP's Statistical review of world energy & EIA's International energy data), Global carbon budget - Fossil CO2 emissions (Global Carbon Project), Global carbon budget - Global carbon emissions (Global Carbon Project), Global carbon budget - National fossil carbon emissions (Global Carbon Project), Global carbon budget - National land-use change carbon emissions (Global Carbon Project), Global carbon budget (Our World in Data based on the Global Carbon Project's Fossil CO2 emissions, Global carbon emissions, National fossil carbon emissions, and National land-use change emissions), Greenhouse gas emissions (including methane and nitrous oxide) by sector (CAIT), CO2 dataset (Our World in Data based on all sources above), Population (Our World in Data based on [a number of different sources](https://ourworldindata.org/population-sources)), and GDP (University of Groningen GGDC's Maddison Project Database, Bolt and van Zanden, 2020). The current dataset we are using extracts from the original dataset.

-   **What processes might have influenced what data was observed and recorded and what was not?**

    -   The current dataset we will use depends on the datasets it was built off of and how they were processed, so if certain countries or emission data were left out of those datasets, they will also be left out of our dataset.

-   **If people are involved, were they aware of the data collection and if so, what purpose did they expect the data to be used for?**

    -   There were no people involved in the dataset.

### Preprocessing/cleaning/labeling

-   **What preprocessing was done, and how did the data come to be in the form that you are using?**

    -   The various datasets that were the foundation for the original dataset underwent processing steps (the codes of which can be found at this website, https://github.com/owid/co2-data), the names of countries and regions were standardized, the carbon emissions were recalculated to CO2 (e.g. from tonnes of carbon to tonnes of CO2 using a conversion factor of 3.664), and per capita figures were calculated from the dataset's "Population" metric, which is included in the complete dataset and the population figures were sourced from Gapminder and the UN World Population Prospects (UNWPP).
    -   The current dataset we will use is a subset of the original dataset, with all the tonnes units converted into kilotons and the variable names altered.

-   **Is the software that was used to preprocess/clean/label the data available?**

    -   The link for the software and code that was used to preprocess/clean/label the data: https://github.com/owid/co2-data

# Data analysis

```{r}
#| label: initial-exploratory-data-analysis

#average GDP of countries
pollution |>
  group_by(Country.Name) |>
  summarize(mean = mean(Country.GDP))

#average total CO2 emissions of countries 
pollution |>
  group_by(Country.Name) |>
  summarize(mean = mean(Emissions.Production.CO2.Total))

#visualizing change in GDP over time for a developed country
pollution_usa <- pollution |>
  filter(Country.Name == "United States")

ggplot(data = pollution_usa, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in USA", 
    x = "Year",
    y = "GDP (international$)"
  ) 

#visualizing change in GDP over time for another developed country
pollution_japan <- pollution |>
  filter(Country.Name == "Japan")

ggplot(data = pollution_japan, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Japan", 
    x = "Year",
    y = "GDP (international$)"
  ) 

#visualizing change in GDP over time for a developing country
pollution_pakistan <- pollution |>
  filter(Country.Name == "Pakistan")

ggplot(data = pollution_pakistan, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Pakistan", 
    x = "Year",
    y = "GDP (international$)"
  ) 

#visualizing change in GDP over time for another developing country
pollution_vietnam <- pollution |>
  filter(Country.Name == "Vietnam")

ggplot(data = pollution_vietnam, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Vietnam", 
    x = "Year",
    y = "GDP (international$)"
  ) 

#visualizing change in CO2 emissions over time for a developed country
ggplot(data = pollution_usa, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in USA", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 

#visualizing change in CO2 emissions over time for another developed country
ggplot(data = pollution_japan, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Japan", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 

#visualizing change in CO2 emissions over time for a developing country
ggplot(data = pollution_pakistan, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Pakistan", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 

#visualizing change in CO2 emissions over time for another developing country
ggplot(data = pollution_vietnam, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Total)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Vietnam", 
    x = "Year",
    y = "Total CO2 Emissions (kilotons)"
  ) 

#visualizing relationship between GDP and total CO2 emissions in 2018
pollution_18 <- pollution |>
  filter(Year == 2018)

ggplot(data = pollution_18,
       mapping = aes(x = Country.GDP, y = Emissions.Production.CO2.Total)
       ) + 
  geom_point(alpha = 0.5) + 
  xlim(0, 4e+12) + 
  ylim(0, 2000) + 
  theme_minimal() + 
  labs(
    title = "Total CO2 Emissions vs. Country GDP in 2018", 
    x = "GDP (international$)",
    y = "Total CO2 Emissions (kilotons)"
  ) 

#visualizing relationship between GDP and total CO2 emissions in 1992
pollution_92 <- pollution |>
  filter(Year == 1992)

ggplot(data = pollution_92,
       mapping = aes(x = Country.GDP, y = Emissions.Production.CO2.Total)
       ) + 
  geom_point(alpha = 0.5) + 
  xlim(0, 3e+12) + 
  ylim(0, 2000) + 
  theme_minimal() + 
  labs(
    title = "Total CO2 Emissions vs. Country GDP in 1992", 
    x = "GDP (international$)",
    y = "Total CO2 Emissions (kilotons)"
  ) 
```

The average GDP for all countries (between 1992 and 2018) in the dataset were calculated and can be found in the displayed tibble.

The average total CO2 emissions for all countries (between 1992 and 2018) in the dataset were calculated and can be found in the second displayed tibble.

Through visualizing the change in GDP over time for the two developed (United States and Japan) countries and the two developing countries (Pakistan and Vietnam), it can be seen that it appears that developing countries have a much higher rate of increase in GDP over time than developed countries do. However, as expected, the developing countries have much lower GDPs than developed countries in general.

Through visualizing the change in total CO2 emissions over time for the two developed (United States and Japan) countries and the two developing countries (Pakistan and Vietnam), it can be seen that it appears that developing countries have a much higher rate of increase in total CO2 emissions over time than developed countries do. Additionally, developed countries' total CO2 emissions don't appear to vary too much over the years, staying at a relatively high emission rate. Furthermore, as expected, despite higher rates of increase in the total CO2 emissions over time for developing countries, developed countries tend to have much higher total CO2 emissions overall than developing countries.

The observations above could indicate a possible correlation between GDP growth and an increase in CO2 emissions. It could also indicate that developed countries tend to produce much more CO2 than developing countries, contributing much more to climate change, but that there is a upward trend in development in developing countries that is also increasing CO2 emissions.

The graphs of "Total CO2 Emissions vs. Country GDP in 2018" and "Total CO2 Emissions vs. Country GDP in 1992" demonstrate a potential positive linear relationship between GDP and total CO2 emissions for the different countries across the years of 1992 to 2018.

# Evaluation of significance

```{r}
#| label: analysis-one

developed_countries <- c(
  "Austria", "Belgium", "Denmark", "Finland", "France", "Germany", "Greece", "Ireland", "Italy", "Luxembourg", "Netherlands", "Portugal", "Spain", "Sweden", "United Kingdom", "Bulgaria", "Croatia", "Cyprus", "Czech Republic", "Estonia", "Hungary", "Latvia", "Lithuania", "Malta", "Poland", "Romania", "Slovakia", "Slovenia", "Iceland", "Norway", "Switzerland", "Australia", "Canada", "Japan", "New Zealand", "United States"
  )
pollution <- pollution |>
  mutate(developed = if_else(Country.Name %in% developed_countries, "Yes", "No"),
         developed = factor(developed))

log_fit <- logistic_reg() |>
  fit(developed ~ Emissions.Production.CO2.Total, data = pollution)

tidy(log_fit)

```

$$
log(\frac{p}{1-p})=-5.59 \times 10^{-1} + 8.20 \times 10^{-5} \times total~CO2~emissions
$$

The intercept can be interpreted as the log odds of a country being "developed" if total CO2 emissions is 0, which is \$-5.59 \\times 10\^{-1}\$. However, because we do not use or think in terms of log odds, we can further interpret this intercept by evaluating p. The odds, $\frac {p}{1-p}$, of a country being "developed" if total CO2 emissions is 0 is $e ^ {-5.59 \times 10 ^ {-1}}$ where p represents the probability that the country is developed. Thus, given that total CO2 emissions is 0, the probability a country is developed is $e ^ {-5.59 \times 10 ^ {-1}} - e ^ {-5.59 \times 10 ^{-1}} p$ .

# Interpretation and conclusions

# Limitations

-   Not all 195 countries are in our dataset, which limits the scope of our dataset a bit.

-   The years within the dataset only range from 1992 to 2018, which could be a limitation because the range is 26 years, which may or may not be a long enough time period to see substantive changes in GDP or CO2 emissions. Additionally, this dataset does not contain data for more recent years, which could be another limitation if we want to see the most up-to-date data on GDP/emissions for different countries.

# Acknowledgments
