---
title: "It's Getting Hot in Here"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: false
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
library(tidymodels)
library(scales)
library(skimr)
```

```{r}
#| label: read-csv-glimpse

pollution <- read_csv("data/global_emissions.csv")
```

# Introduction

Climate change is the biggest challenge facing our generation, with increased global CO2 emissions being one of the leading contributors to this issue. According to NASA, "Carbon dioxide in the atmosphere warms the planet, causing climate change. Human activities have raised the atmosphere's carbon dioxide content by 50% in less than 200 years," and according to Climate.gov, " By adding more carbon dioxide to the atmosphere, people are supercharging the natural greenhouse effect, causing global temperature to rise." However, in order to combat the effects of climate change, we must first understand the data behind it. In our analysis, we intend to answer the following research questions:

1.  In 2018, is the true proportion of developed countries with CO2 gas emissions above 1000 kilotons different than those with CO2 gas emissions below 1000 kilotons? The definition of developing versus devleoped countries will be as defined by the United Nations:

    https://www.un.org/en/development/desa/policy/wesp/wesp_current/2014wesp_country_classification.pdf?

2.  Is the true proportion of countries with a CO2 emission greater than the US in the twentieth century different from those in the twenty-first century?

*\*\*ADD BRIEF SUMMARY OF RESULTS*

# Data description

### Motivation

The dataset was created to show how emissions are changing in each country and allows for comparisons of various countries and regions in relation to overall emission trends along with their sources. The Our World in Data team created the original dataset on behalf of the Our World in Data project, which is a project of the Global Change Data Lab (a non-profit organization based in the United Kingdom), and the current dataset we will be using is a subset from the original dataset from the CORGIS Dataset Project created by Sam Donald. The Global Change Data Lab funded the creation of the dataset and it is a registered charity in England and Wales. Their work is made possible by donations of individuals and organizations that have made contributions to support their project.

### Composition

The instances that comprise the dataset represent countries (per year from 1992 to 2018) and there are 2484 instances total. The observations (rows) are countries per year and the attributes (columns) are year, country name, country code, country GDP, country population, emissions production CH4, emissions production N2O, emissions production CO2 cement, emissions production CO2 coal, emissions production CO2 gas, emissions production CO2 oil, emissions production CO2 flaring, emissions production CO2 other, emissions production CO2 total, emissions global share CO2 cement, emissions global share CO2 coal, emissions global share CO2 gas, emissions global share CO2 oil, emissions global share CO2 flaring, emissions global share CO2 total.

### Collection Process

The original dataset is built upon a number of datasets and processing steps. The datasets that were processed include Statistical review of world energy (BP), International energy data (EIA), Primary energy consumption (Our World in Data based on BP's Statistical review of world energy & EIA's International energy data), Global carbon budget - Fossil CO2 emissions (Global Carbon Project), Global carbon budget - Global carbon emissions (Global Carbon Project), Global carbon budget - National fossil carbon emissions (Global Carbon Project), Global carbon budget - National land-use change carbon emissions (Global Carbon Project), Global carbon budget (Our World in Data based on the Global Carbon Project's Fossil CO2 emissions, Global carbon emissions, National fossil carbon emissions, and National land-use change emissions), Greenhouse gas emissions (including methane and nitrous oxide) by sector (CAIT), CO2 dataset (Our World in Data based on all sources above), Population (Our World in Data based on [a number of different sources](https://ourworldindata.org/population-sources)), and GDP (University of Groningen GGDC's Maddison Project Database, Bolt and van Zanden, 2020). The current dataset we are using extracts from the original dataset.

The current dataset we will use depends on the datasets it was built off of and how they were processed, so if certain countries or emission data were left out of those datasets, they will also be left out of our dataset.

### Preprocessing/cleaning/labeling

The various datasets that were the foundation for the original dataset underwent processing steps (the codes of which can be found at this website, https://github.com/owid/co2-data), the names of countries and regions were standardized, the carbon emissions were recalculated to CO2 (e.g. from tonnes of carbon to tonnes of CO2 using a conversion factor of 3.664), and per capita figures were calculated from the dataset's "Population" metric, which is included in the complete dataset and the population figures were sourced from Gapminder and the UN World Population Prospects (UNWPP). The current dataset we will use is a subset of the original dataset, with all the tonnes units converted into kilotons and the variable names altere

The link for the software and code that was used to preprocess/clean/label the data: https://github.com/owid/co2-data

# Data analysis

```{r}
#| label: eda-avg-gdp

#average GDP of countries
pollution |>
  group_by(Country.Name) |>
  summarize(mean = mean(Country.GDP))

```

The average GDP for all countries (between 1992 and 2018) in the dataset were calculated and can be found in the displayed tibble.

```{r}
#| label: eda-avg-co2

#average total CO2 emissions of countries 
pollution |>
  group_by(Country.Name) |>
  summarize(mean = mean(Emissions.Production.CO2.Gas))
```

The average CO2 gas emissions for all countries (between 1992 and 2018) in the dataset were calculated and can be found in the second displayed tibble.

```{r}
#| label: eda-gdp-developed

#visualizing change in GDP over time for a developed country
pollution_usa <- pollution |>
  filter(Country.Name == "United States")

ggplot(data = pollution_usa, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in USA", 
    x = "Year",
    y = "GDP (international$)"
  ) 

#visualizing change in GDP over time for another developed country
pollution_japan <- pollution |>
  filter(Country.Name == "Japan")

ggplot(data = pollution_japan, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Japan", 
    x = "Year",
    y = "GDP (international$)"
  ) 
```

```{r}
#| label: eda-gdp-developing

#visualizing change in GDP over time for a developing country
pollution_pakistan <- pollution |>
  filter(Country.Name == "Pakistan")

ggplot(data = pollution_pakistan, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Pakistan", 
    x = "Year",
    y = "GDP (international$)"
  ) 

#visualizing change in GDP over time for another developing country
pollution_vietnam <- pollution |>
  filter(Country.Name == "Vietnam")

ggplot(data = pollution_vietnam, mapping = aes(x = Year, y = Country.GDP)) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "GDP Change Over Time in Vietnam", 
    x = "Year",
    y = "GDP (international$)"
  ) 
```

Through visualizing the change in GDP over time for the two developed (United States and Japan) countries and the two developing countries (Pakistan and Vietnam), it can be seen that it appears that developing countries have a much higher rate of increase in GDP over time than developed countries do. However, as expected, the developing countries have much lower GDPs than developed countries in general.

```{r}
#| label: eda-co2-developed

#visualizing change in CO2 emissions over time for a developed country
ggplot(data = pollution_usa, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Gas)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in USA", 
    x = "Year",
    y = "CO2 Gas Emissions (kilotons)"
  ) 

#visualizing change in CO2 emissions over time for another developed country
ggplot(data = pollution_japan, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Gas)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Japan", 
    x = "Year",
    y = "CO2 Gas Emissions (kilotons)"
  ) 

```

```{r}
#| label: eda-co2-developing

#visualizing change in CO2 emissions over time for a developing country
ggplot(data = pollution_pakistan, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Gas)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Pakistan", 
    x = "Year",
    y = "CO2 Gas Emissions (kilotons)"
  ) 

#visualizing change in CO2 emissions over time for another developing country
ggplot(data = pollution_vietnam, 
       mapping = aes(x = Year, y = Emissions.Production.CO2.Gas)
       ) +
  geom_col() + 
  theme_minimal() +
  labs(
    title = "CO2 Emissions Change Over Time in Vietnam", 
    x = "Year",
    y = "CO2 Gas Emissions (kilotons)"
  ) 
```

Through visualizing the change in CO2 gas emissions over time for the two developed (United States and Japan) countries and the two developing countries (Pakistan and Vietnam), it can be seen that it appears that developing countries have a much higher rate of increase in CO2 gas emissions over time than developed countries do. Furthermore, as expected, despite higher rates of increase in the CO2 gas emissions over time for developing countries, developed countries tend to have much higher CO2 gas emissions overall than developing countries.

The observations above could indicate a possible correlation between GDP growth and an increase in CO2 gas emissions. It could also indicate that developed countries tend to produce much more CO2 gas than developing countries, contributing much more to climate change, but that there is a upward trend in development in developing countries that is also increasing CO2 gas emissions.

To analyze the relationship between GDP and CO2 gas emissions for developing versus developed countries in 2018, we will utilize a logistic regression model. As there is no clear definition of a GDP threshold that determines whether a country is developed or developing, we will use the the United Nations' definition (https://www.un.org/en/development/desa/policy/wesp/wesp_current/2014wesp_country_classification.pdf?), since their definition takes a country's GDP into account.

```{r}
#| label: eda-2018-gdp-co2

developed_countries <- c(
  "Austria", "Belgium", "Denmark", "Finland", "France", "Germany", "Greece", "Ireland", "Italy", "Luxembourg", "Netherlands", "Portugal", "Spain", "Sweden", "United Kingdom", "Bulgaria", "Croatia", "Cyprus", "Czech Republic", "Estonia", "Hungary", "Latvia", "Lithuania", "Malta", "Poland", "Romania", "Slovakia", "Slovenia", "Iceland", "Norway", "Switzerland", "Australia", "Canada", "Japan", "New Zealand", "United States"
  )

pollution <- pollution |>
  mutate(
    developed = if_else(Country.Name %in% developed_countries, "developed", "developing"),
         developed = factor(developed))

pollution_18 <- pollution |>
  filter(Year == 2018)

log_fit <- logistic_reg() |>
  fit(developed ~ Emissions.Production.CO2.Gas, data = pollution_18)

tidy(log_fit)

```

The logistic regression model using the total CO2 emissions production as an explanatory variable is as follows:

$$
log(\frac{p}{1-p})= 0.573 - 0.000483 \times CO2~gas~emissions
$$

The probability a country is developed can be calculated as:

$$p = \frac{\exp\{0.573 - 0.000483 \times CO2~gas~emissions\}}{1 + \exp\{0.573 - 0.000483 \times CO2~gas~emissions\}}$$

Thus, using the logistic regression model, if the CO2 gas emissions of a given country is 100 kilotons, the probability that that country is developed is 0.37:

```{r}
#| label: pred-log-odds-1

co2_100 <- tibble(Emissions.Production.CO2.Gas = 100)
predict(log_fit, new_data = co2_100, type = "prob")
```

If the CO2 gas emissions of a given country is 1000 kilotons, the probability that that country is developed is 0.48:

```{r}
#| label: pred-log-odds-2

co2_1000 <- tibble(Emissions.Production.CO2.Gas = 1000)
predict(log_fit, new_data = co2_1000, type = "prob")
```

If the CO2 gas emissions of a given country is 3000 kilotons, the probability that that country is developed is 0.71:

```{r}
#| label: pred-log-odds-3

co2_3000 <- tibble(Emissions.Production.CO2.Gas = 3000)
predict(log_fit, new_data = co2_3000, type = "prob")
```

Using the default decision-making threshold of 0.5 (that is, if the predicted probability that a country is developed is greater than 0.5, then we would classify it as developed), we can see that as the CO2 gas emissions increases, the probability that the country is developed also increases, which indicates that GDP growth/higher GDP is correlated with higher CO2 gas emissions. Additionally, it appears that there is a CO2 gas emission threshold of around 1000 kilotons, where countries with CO2 gas emissions greater than 1000 kilotons are most likely developed, while countries with CO2 gas emissions less than 1000 kilotons are most likely developing.

# Evaluation of significance

## Research Question One

The first research question we will attempt to answer is: In 2018, is the true proportion of developed countries with CO2 gas emissions above 1000 kilotons different than those with CO2 gas emissions below 1000 kilotons? As there is no clear definition of a GDP threshold that determines whether a country is developed or developing, we will once again use the the United Nations' definition (https://www.un.org/en/development/desa/policy/wesp/wesp_current/2014wesp_country_classification.pdf?), since their definition takes a country's GDP into account. To attempt to answer this question, we will carry out a hypothesis test.

Using the observed potential CO2 gas emission threshold of around 1000 kilotons from the logistic regression model, the null hypothesis is that in 2018, the true proportion of developed countries with CO2 gas emissions above 1000 kilotons is not different than those with CO2 gas emissions below 1000 kilotons, and the alternative hypothesis is that in 2018, the true proportion of developed countries with CO2 gas emissions above 1000 kilotons is different than those with CO2 gas emissions below 1000 kilotons. In proper notation:

$$
H_0 : p_a - p_b = 0
$$ $$
H_A: p_a - p_b \neq 0
$$

```{r}
#| label: research-question-one

set.seed(123)

pollution_18 <- pollution_18 |>
  mutate(
    co2_above_1000 = if_else(Emissions.Production.CO2.Gas > 1000, "yes", "no"), 
    co2_above_1000 = parse_factor(co2_above_1000)
  )

#calculate point estimate
point_estimate <- pollution_18 |>
  specify(developed ~ co2_above_1000, success = "developed") |>
  calculate(stat = "diff in props", order = c("yes", "no"))

#create null distribution
null_dist <- pollution_18 |>
  specify(developed ~ co2_above_1000, success = "developed") |>
  hypothesize(null = "independence") |>
  generate(1000, type = "permute") |>
  calculate(stat = "diff in props", order = c("yes", "no"))

#calculate p-value
null_dist |>
  get_p_value(obs_stat = point_estimate, direction = "two sided")

```

Since the p-value is 0.766, which is greater than the significance level of 0.05, we fail to reject the null hypothesis. The data fails to provide convincing evidence that in 2018, the proportion of developed countries with CO2 gas emissions above 1000 kilotons is different than those with CO2 gas emissions below 1000 kilotons.

## Research Question Two

The second research question we will attempt to answer is if the true proportion of countries with a CO2 emission greater than the US in the twentieth century different from those in the twenty-first century. We will do this using a null hypothesis test. The null hypothesis is that there is no difference between the centuries, and the alternative hypothesis is the opposite: that there is a difference. In proper notation:

$$
H_0 : p_a - p_b = 0
$$ $$
H_A: p_a - p_b \neq 0
$$

```{r}
#| label: research-question-two

set.seed(123)

us_2000 <- as.integer(pollution[pollution$Country.Name == "United States" & pollution$Year == 2000, "Emissions.Production.CO2.Total"])


pollution_mod <- pollution |>
  mutate(
    us_compare = if_else(Emissions.Production.CO2.Total > us_2000, "more" , "less"), 
    us_compare = as.factor(us_compare),
    y2k = if_else(Year >= 2000, "twentyfirst", "twentieth"),
    y2k = as.factor(y2k)
  ) 

point_estimate <- pollution_mod |>
  specify(us_compare ~ y2k, success = "more") |>
  calculate(stat = "diff in props", order = c("twentieth", "twentyfirst"))

null_dist <- pollution_mod |>
  specify(us_compare ~ y2k, success = "more") |>
  hypothesize(null = "independence") |>
  generate(1000, type = "permute") |>
  calculate(stat = "diff in props", order = c("twentieth", "twentyfirst"))

visualize(null_dist) +
 shade_p_value(obs_stat = point_estimate, direction = "two sided")

null_dist |>
  get_p_value(obs_stat = point_estimate, direction = "two sided")
```

Since the p-value is 0, which is less than the significance level of 0.05, we reject the null hypothesis in favor of the alternative hypothesis. The data provides convincing evidence that the true proportion of countries with a CO2 emission greater than the US in the twentieth century is different from those in the twenty-first century.

# Interpretation and conclusions

We find that there is indeed an increase in CO2 emissions, and that this increase is correlated to GDP of a country. Our data indicates a dramatic increase in CO2 emissions; for developing countries, a near-exponential increase, and for developed countries, a linear increase.

Regardless, according to the Intergovernmental Panel on Climate Change (IPCC), the United Nations Environment Program (UNEP), and the United States Environmental Protection Agency (EPA), Carbon dioxide emissions are problematic because they contribute to the warming of the Earth's atmosphere and the subsequent climate change.

In developing nations, we tend to see industrialization, which means more burning of fossil fuels and deforestation, thus increasing CO2 emissions. Thus, we see increase in emission of these countries.

Reducing emissions is important to mitigate the impacts of climate change and limit the ever-increasing global temperature. This requires transitioning to renewable energy sources, improving energy efficiency, reducing deforestation, and promoting reforestation and eco-repair. Developing or developed, every nation has a duty to help protect and preserve the planet in which we all reside.

Start off with C02 increase is steerper in devleoped than developing. (summarize what's in EDA) Which developes our first question, which turns out to be not significant.

THen second

And conclude

# Limitations

-   Not all 195 countries are in our dataset, which limits the scope of our dataset a bit.

-   The years within the dataset only range from 1992 to 2018, which could be a limitation because the range is 26 years, which may or may not be a long enough time period to see substantive changes in GDP or CO2 emissions. Additionally, this dataset does not contain data for more recent years, which could be another limitation if we want to see the most up-to-date data on GDP/emissions for different countries.

# Acknowledgments
